---
title: "Soil Moisture Measurements from Sentinel-2 Data"
graphics: yes
output: html_notebook
---

```{r knitr setup, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
```

We are going to be pulling Sentinel-2 imagery from Google Cloud, using the Google Cloud SDK (CLI). In order for the rest of this process to work, make sure that [Google Cloud SDK](https://cloud.google.com/sdk/docs/install) is installed on your computer in the default directory.

```{r setup packages, warning=FALSE, message=FALSE}
library(sen2r) #read in sentinel-2 imagery
library(h3) #aggrigation on Ubers H3 hexagon grid
library(terra) #Spatial data tools
library(leaflet) #Interactive map viewer
library(sf) #for conversion between spatial formats

check_gcloud() 
```

If you get an error about Google Cloud Initialization make sure you open a command prompt (cmd) and enter `gcloud init` .

# Create Extent

First we will need to create a feature that encompasses the area of interest, in this case - Laguna Lake in San Luis Obispo, CA. First we use an online map to get some corner coordinates and create a spatial vector. Then we make sure it's in the right place by plotting them on a leaflet map.

```{r Laguna Extent}

Laguna_Ext <- vect("POLYGON ((-120.7063985 35.2706786, -120.6881986 35.2813860, -120.6709197 35.2768445, -120.6741943 35.2647444,-120.6883474 35.2544259, -120.7063985 35.2706786))")

leaflet(st_as_sf(Laguna_Ext)) %>% 
  addTiles() %>% 
  addPolygons()
```

# Download Sentinel-2 Data

```{r}

sen2r()
```
